### 2016-02-29 Manuscript/SW_early_life_telomeres_tables_and_figures.Rmd
More major edits

Fig1A <- ggplot(temp,aes(x = Agemonths,y = LogTL))+
  geom_jitter(col='grey') +
  ylab('Log telomere length (kb)') + 
  xlab('Age (months)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'none')+
  geom_boxplot(outlier.colour=NA)+
  annotate('text',x = 0.6,y = 4.8,label = 'A')

#
temp <- Loss
temp[,'Age (months)'] <- factor(temp$Agemonths)

Fig1B <- ggplot(temp,aes(x = LogTL,y = LogTL1,col = `Age (months)`,shape = `Age (months)`)) +
  ylab('Log adult telomere length (kb)') + 
  xlab('Log juvenile telomere length (kb)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'none')+
  geom_point() +
  stat_smooth(method = 'lm')+
  xlim(2,4.5)+
  ylim(2,4.5)+
    scale_colour_manual(values = c('darkgrey','darkgreen','darkorange'))+
  annotate('text',x = 2,y = 4.5,label = 'B')

#
temp <- Loss
temp[,'Age (months)'] <- factor(temp$Agemonths)

Fig1C <- ggplot(temp,aes(x = TimeDiff,y = Loss,col=`Age (months)`,shape = `Age (months)`)) +
  ylab('Log telomere loss (kb)') + 
  xlab('Time interval (years)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'none')+
  geom_point()+
  geom_abline(slope = 0,intercept = 0,lty = 2)+
  stat_smooth(method='lm')+
  scale_colour_manual(values = c('darkgrey','darkgreen','darkorange'))+
  annotate('text',x = 0,y = 2,label = 'C')

Fig1D <- ggplot(temp,aes(x = TimeDiff,y = TROC,col=`Age (months)`,shape = `Age (months)`)) +
  ylab('Log telomere loss rate (kb/year)') + 
  xlab('Time interval (years)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = c(0.9,0.8))+
  geom_point()+
  geom_abline(slope = 0,intercept = 0,lty = 2)+
  scale_colour_manual(values = c('darkgrey','darkgreen','darkorange'))+
  annotate('text',x = 0,y = 8,label = 'D')

pdf('Figures/Figure 1.pdf',
    width = 8,
    height = 8,
    paper = 'a4r')
multiplot(Fig1A,Fig1B,Fig1C,Fig1D,layout=matrix(c(1:4),nrow = 2,byrow=T))
dev.off()

@@

###############################
temp <- juv
temp$LayYear <- factor(temp$LayYear)

Fig2A <- ggplot(subset(temp,Agemonths == 1),
       aes(x = LayYear,
           y = LogTL)) +
  ylab('') + 
  xlab('') +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 60,size = 14,vjust=0.5), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_jitter(col = 'grey')+
  ylim(c(2,4.5))+
  geom_boxplot(outlier.colour = NA)+
  scale_x_discrete(limits = factor(1998:2014),drop=F)+
  annotate('text',x = 1,y = 4.4,label = 'A')

Fig2B <- ggplot(subset(juv,Agemonths == 6),
       aes(x = factor(LayYear),
           y = LogTL)) +
  ylab('Log telomere length (kb)') + 
  xlab('') +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 60,size = 14,vjust=0.5), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_jitter(col = 'grey')+
  ylim(c(2,4.5))+
  geom_boxplot(outlier.colour = NA)+
  scale_x_discrete(limits = factor(1998:2014),drop=F)+
  annotate('text',x = 1,y = 4.4,label = 'B')

Fig2C <- ggplot(subset(juv,Agemonths == 10),
       aes(x = factor(LayYear),
           y = LogTL)) +
  ylab('') + 
  xlab('Cohort') +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 60,size = 14,vjust=0.5), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_jitter(col = 'grey')+
  ylim(c(2,4.5))+
  geom_boxplot(outlier.colour = NA)+
  scale_x_discrete(limits = factor(1998:2014),drop=F)+
  annotate('text',x = 1,y = 4.4,label = 'C')

pdf('Figures/Figure 2.pdf',
    width = 6,
    height = 8,
    paper = 'a4r')
multiplot(Fig2A,Fig2B,Fig2C)
dev.off()

@@

#Run the model
lm2 <- lmer(LogTL ~ Agemonths * (Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ)  + (1|LayYear) + (1|PlateID) + (1|TerritoryID),
            data = temp,
            REML = FALSE)
lm2 <- standardize(lm2)
#Model averaging

@@

Fig3 <- forestplot(table2)+
  ylim(-0.4,0.4)

pdf('Figures/Figure 3.pdf',
    width = 8,
    height = 5,
    paper = 'a4r')
Fig3
dev.off()

@@

```{r Figure 4}
temp <- juv

@@

Fig4A <- ggplot(temp,
       aes(x = factor(Agemonths),
           y = LogTL,
           fill=`Non helpers`)) +
  ylab('Log telomere length (kb)') + 
  xlab('Age (months)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_jitter(aes(colour = `Non helpers`),position=position_jitterdodge(dodge.width = 0.8))+
  geom_boxplot(outlier.colour = NA,notch=F)+
  scale_colour_manual(values = c('grey','darkorange'))+
  scale_fill_manual(values = c('grey','darkorange'))+
  annotate('text',x = 0.55,y = 4.5,label = 'A')

Fig4B <- ggplot(temp,
       aes(x = factor(Agemonths),
           y = LogTL,
           fill=`Helpers`)) +
  ylab('Log telomere length (kb)') + 
  xlab('Age (months)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_jitter(aes(colour = `Helpers`),position=position_jitterdodge(dodge.width = 0.8))+
  geom_boxplot(outlier.colour = NA,notch=F)+
  scale_colour_manual(values = c('grey','darkorange'))+
  scale_fill_manual(values = c('grey','darkorange'))+
  annotate('text',x = 0.55,y = 4.5,label = 'B')

Fig4C <- ggplot(temp,
       aes(x = TQ,
           y = LogTL,
           colour=`Age (months)`,
           fill = `Age (months)`,
           shape = `Age (months)`)) +
  ylab('Log telomere length (kb)') + 
  xlab('Territory quality') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_point(alpha = 0.2)+
  stat_smooth(method = 'lm')+
  scale_colour_manual(values = c(grey(0.2),'darkorange','darkgreen'))+
  scale_fill_manual(values = c('darkgrey','darkorange','darkgreen'))+
  annotate('text',x = 3,y = 4.5,label = 'C')

Fig4D <- ggplot(temp,
       aes(x = Tarsus,
           y = LogTL,
           colour=`Age (months)`,
           fill = `Age (months)`,
           shape = `Age (months)`)) +
  ylab('Log telomere length (kb)') + 
  xlab('Tarsus length (mm)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_point(alpha = 0.2)+
  stat_smooth(method = 'lm')+
  scale_colour_manual(values = c(grey(0.2),'darkorange','darkgreen'))+
  scale_fill_manual(values = c('darkgrey','darkorange','darkgreen'))+
  annotate('text',x = 17,y = 4.5,label = 'D')

pdf('Figures/Figure 4.pdf',
    width = 7,
    height = 8,
    paper = 'a4r')
multiplot(Fig4A,Fig4B,Fig4C,Fig4D,layout=matrix(c(1:4),byrow = TRUE,nrow = 2))
dev.off()

@@

```{r Figure 5}
temp <- subset(juv,LayYear<2009)
mymed <- median(temp$cohortTL)
temp$coTLF <- ifelse(temp$cohortTL>mymed,'h','l')

s1 <- survfit(Surv(time = Lifespan, event = Died)~coTLF, data = temp)
Fig5A <- ggsurv(s1,plot.cens=F,lty.est=c(1:2)) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  xlim(c(0,15)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  scale_colour_manual(name = 'TL',
                      values = c('grey','black')) +
  guides(colour=F)

mymed <- median(temp$cenTL)
temp$cenTLF <- ifelse(temp$cenTL>mymed,'h','l')
s1 <- survfit(Surv(time = Lifespan, event = Died)~cenTLF, data = temp)
Fig5B <- ggsurv(s1,plot.cens=F,lty.est=c(1,2)) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  xlim(c(0,15)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  scale_colour_manual(name = 'TL',
                      values = c('black','grey')) +
  guides(colour=F)

pdf('Figures/Figure 5.pdf',
    width = 4,
    height = 8,
    paper = 'a4')
multiplot(Fig5A,Fig5B)
dev.off()

>>>>

Fig1A <- ggplot(temp,aes(x = Agemonths,y = RTL))+
  geom_jitter(col='grey') +
  ylab('RTL') + 
  xlab('Age (months)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'none')+
  geom_boxplot(outlier.colour=NA)
  annotate('text',x = 0.6,y = 4.8,label = 'A')

  temp <- Loss
  temp[,'Age (months)'] <- factor(temp$Agemonths)

Fig1B <- ggplot(temp,aes(x = `Age (months)`,y = TROC)) +
    geom_jitter(col='grey') +
  ylab('Loss in RTL') + 
  xlab('Age (months)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = c(0.9,0.8))+
  geom_boxplot(outlier.shape = NA)+
  stat_smooth(method = 'lm', formula = y~(1/exp(x)))+
  scale_colour_manual(values = c('darkgrey','darkgreen','darkorange'))
  annotate('text',x = 0,y = 8,label = 'D')

temp <- juv
temp$LayYear <- factor(temp$LayYear)

Fig1C <- ggplot(temp,
       aes(x = LayYear,
           y = RTL)) +
  ylab('') + 
  xlab('RTL') +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 60,size = 14,vjust=0.5), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_jitter(col = 'grey')+
  geom_boxplot(outlier.colour = NA)+
  scale_x_discrete(limits = factor(1998:2014),drop=F)
  annotate('text',x = 1,y = 4.4,label = 'C')

temp <- Loss
temp$LayYear <- factor(temp$LayYear)

Fig1D <- ggplot(temp,
       aes(x = LayYear,
           y = RTL)) +
  ylab('') + 
  xlab('Loss in RTL') +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 60,size = 14,vjust=0.5), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_jitter(col = 'grey')+
  geom_boxplot(outlier.colour = NA)+
  scale_x_discrete(limits = factor(1998:2014),drop=F)
  annotate('text',x = 1,y = 4.4,label = 'D')

pdf('Figures/Figure 1.pdf',
    width = 8,
    height = 8,
    paper = 'a4r')
multiplot(Fig1A,Fig1B,Fig1C,Fig1D,layout=matrix(c(1,2,3,3,4,4),nrow = 3,byrow=T))
dev.off()

@@

@@

#Run the model
lm2 <- lmer(RTL ~ Agemonths * (Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ)  + (1|LayYear) + (1|TerritoryID),
            data = temp,
            REML = FALSE)
lm2 <- standardize(lm2)
#Model averaging

@@

Fig2A <- forestplot(table2)

Fig2B <- ggplot(temp,
       aes(x = factor(Agemonths),
           y = RTL,
           fill=`Helpers`)) +
  ylab('RTL') + 
  xlab('Age (months)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_jitter(aes(colour = `Helpers`),position=position_jitterdodge(dodge.width = 0.8))+
  geom_boxplot(outlier.colour = NA,notch=F)+
  scale_colour_manual(values = c('grey','darkorange'))+
  scale_fill_manual(values = c('grey','darkorange'))
  annotate('text',x = 0.55,y = 4.5,label = 'B')

Fig2C <- ggplot(temp,
       aes(x = TQ,
           y = RTL,
           colour=`Age (months)`,
           fill = `Age (months)`,
           shape = `Age (months)`)) +
  ylab('RTL') + 
  xlab('Territory quality') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_point(alpha = 0.2)+
  stat_smooth(method = 'lm')+
  scale_colour_manual(values = c(grey(0.2),'darkorange','darkgreen'))+
  scale_fill_manual(values = c('darkgrey','darkorange','darkgreen'))
  annotate('text',x = 3,y = 4.5,label = 'C')

Fig2D <- ggplot(temp,
       aes(x = Tarsus,
           y = RTL,
           colour=`Age (months)`,
           fill = `Age (months)`,
           shape = `Age (months)`)) +
  ylab('RTL') + 
  xlab('Tarsus length (mm)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_point(alpha = 0.2)+
  stat_smooth(method = 'lm')+
  scale_colour_manual(values = c(grey(0.2),'darkorange','darkgreen'))+
  scale_fill_manual(values = c('darkgrey','darkorange','darkgreen'))
  annotate('text',x = 17,y = 4.5,label = 'D')

pdf('Figures/Figure 2.pdf',
    width = 9,
    height = 9,
    paper = 'a4r')
multiplot(Fig2A,Fig2B,Fig2C,Fig2D,layout = matrix(c(1,1,1,2,3,4),nrow = 2,byrow = T))
dev.off()

@@

temp <- subset(Loss,TimeDiff>0.3)

#Run the model
lm2 <- lmer(TROC ~ Agemonths * (Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ)  + RTL + (1|LayYear),
            data = temp,
            REML = FALSE)
lm2 <- standardize(lm2)
#Model averaging

#Sort out row/col names
nams <- rownames(table2)
nams[nams == 'z.RTL'] <- 'RTL'
nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex'
nams[nams == 'z.TQ'] <- 'TQ'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.Agemonths'] <- 'Age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.Agemonths:z.NonHelper'] <- 'Age x non helpers'
nams[nams == 'c.Sex:z.Agemonths'] <- 'Age x sex'
nams[nams == 'z.Agemonths:z.TQ'] <- 'Age x TQ'
nams[nams == 'c.Helper:z.Agemonths'] <- 'Age x helpers'
nams[nams == 'z.Agemonths:z.BodyMass'] <- 'Age x mass'
nams[nams == 'z.Agemonths:z.Tarsus'] <- 'Age x tarsus'

Fig3A <- forestplot(table2)

@@

Fig3B <- ggplot(temp,
       aes(x = factor(Agemonths),
           y = TROC,
           fill=`Helpers`)) +
  ylab('Loss in RTL') + 
  xlab('Age (months)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_jitter(aes(colour = `Helpers`),position=position_jitterdodge(dodge.width = 0.8))+
  geom_boxplot(outlier.colour = NA,notch=F)+
  scale_colour_manual(values = c('grey','darkorange'))+
  scale_fill_manual(values = c('grey','darkorange'))
  annotate('text',x = 0.55,y = 4.5,label = 'B')

Fig3C <- ggplot(temp,
       aes(x = factor(Agemonths),
           y = TROC,
           fill=Sex)) +
  ylab('Loss in RTL') + 
  xlab('Age (months)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_jitter(aes(colour = Sex),position=position_jitterdodge(dodge.width = 0.8))+
  geom_boxplot(outlier.colour = NA,notch=F)+
  scale_colour_manual(values = c('grey','darkorange'))+
  scale_fill_manual(values = c('grey','darkorange'))
  annotate('text',x = 0.55,y = 4.5,label = 'B')

Fig3D <- ggplot(temp,
       aes(x = BodyMass,
           y = TROC,
           col=`Age (months)`)) +
  ylab('Loss in RTL') + 
  xlab('Body mass (g)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_point(alpha = 0.2)+
  stat_smooth(method = 'lm')+
    scale_colour_manual(values = c(grey(0.2),'darkorange','darkgreen'))+
  scale_fill_manual(values = c('darkgrey','darkorange','darkgreen'))
  annotate('text',x = 17,y = 4.5,label = 'D')

pdf('Figures/Figure 3.pdf',
    width = 9,
    height = 9,
    paper = 'a4r')
multiplot(Fig3A,Fig3B,Fig3C,Fig3D,layout = matrix(c(1,1,1,2,3,4),nrow = 2,byrow = T))
dev.off()

@@

```{r Figure 4}
temp <- subset(juv,LayYear<2009)
mymed <- median(temp$cohortTL)
temp$coTLF <- ifelse(temp$cohortTL>mymed,'h','l')

s1 <- survfit(Surv(time = Lifespan, event = Died)~coTLF, data = temp)
Fig4A <- ggsurv(s1,plot.cens=F,lty.est=c(1:2)) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  xlim(c(0,15)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  scale_colour_manual(name = 'TL',
                      values = c('grey','black')) +
  guides(colour=F)

temp <- subset(temp,!(is.na(cenTL)))
mymed <- median(temp$cenTL)
temp$cenTLF <- ifelse(temp$cenTL>mymed,'h','l')

s1 <- survfit(Surv(time = Lifespan, event = Died)~cenTLF, data = temp)
Fig4B <- ggsurv(s1,plot.cens=F,lty.est=c(1,2)) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  xlim(c(0,15)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  scale_colour_manual(name = 'TL',
                      values = c('black','grey')) +
  guides(colour=F)

pdf('Figures/Figure 4.pdf',
    width = 4,
    height = 8,
    paper = 'a4')
multiplot(Fig4A,Fig4B)
dev.off()
